---
title: "anonymization"
output: pdf_document
date: "2024-11-05"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
require(dplyr)  
require(sdcMicro)
require(readxl)
require(eeptools)
require(openxlsx)
require(xlsx)
require(plyr)

```

## R Markdown

Step 2: Anonymise the raw dataset and calculate the disclosure risks (you can use any suitable metrics, e.g. among these that were discussed in the course or presented in optional readings). In these calculations, assume that your anonymised dataset, the public data (that is, the files public_data_registerX.xslx and public_data_resultsX.xslx) as well as your submitted report (see Step 4 below) will be available to a potential adversary, but no other data will be available.

```{r}
pri_data <- read_excel("data/private_dataE.xlsx", sheet =1)
# pub_reg_data <- read_excel("public_data_registerE.xlsx")
# pub_res_data <- read_excel("public_data_resultsE.xlsx")
#print(typeof(pri_data))

pri_data$dob <- as.Date(pri_data$dob)
pri_data$age <- floor(age_calc(pri_data$dob, units = "years"))

#pri_data_test <- head(pri_data,10)

print(pri_data)
#print(typeof(pri_data))
#print(typeof(df))

```
Creating a sdc object and computing initial risks: global risk and k-anonymity 
```{r}
sdc_obj <- createSdcObj(dat = pri_data, keyVars = c("dob", "sex", "citizenship", "education", "marital_status", "zip"))
#print(sdc_obj@risk)
print(sdc_obj, type="kAnon")
print(sdc_obj, "risk")
# Global risk (average re-identification probability)
print(sdc_obj@risk$global$risk)

#print(sdc_obj)
# Print L-Diversity
 # Output for l-diversity
#sdc_obj@risk$global

```
Removing direct identifiers: Name
```{r}
pri_data <- subset(pri_data, select = -name)
```

Global recoding on age -> age groups 
Pram on sex variable 
```{r}
#globalRecode(pri_data$age, breaks = 6)
pri_data_pram_age <- pri_data
pri_data_pram_age$age_group <- cut(pri_data_pram_age$age, breaks=c(18, 30, 38, 48, 58, 70, 101), 
                    right=TRUE)

# creating object 
sdc_obj <- createSdcObj(dat = pri_data_pram_age, keyVars = c("age_group", "sex", "citizenship", "education", "marital_status", "zip"))

sdc_obj_k <- createSdcObj(dat = pri_data_pram_age, keyVars = c("age_group", "citizenship", "marital_status", "zip"))
print(sdc_obj_k, type="kAnon")
print(sdc_obj_k, "risk")

# pram on sex variable 
sdc_obj <- pram(sdc_obj, variables = "sex", strata_variables = NULL, pd = 0.7, alpha = 0.5)

#key_vars_subset <- which(names(pri_data) %in% c("age_group", "citizenship", "marital_status", "zip"))
#sdc_obj@keyVars <- key_vars_subset

#print(sdc_obj@keyVars)
#sdc_obj@keyVars <- c("age_group", "citizenship", "marital_status", "zip")

#manip_data <- get.sdcMicroObj(sdc_obj, type="manipKeyVars")

anonymized_data <- extractManipData(sdc_obj)

#print(anonymized_data)

anon_data <- anonymized_data[, !(colnames(anonymized_data) %in% c("dob", "age"))]
print(anon_data)

#write.xlsx(anon_data, file = "anon_data/anon_data_global_recoding_pram.xlsx")

```
Global recoding on age 
Global recoding on marital status
Pram on sex
```{r}
#global recode on age 
pri_data_pram_age_mar <- pri_data
pri_data_pram_age_mar$age_group <- cut(pri_data_pram_age_mar$age, breaks=c(18, 30, 38, 48, 58, 70, 101), 
                    right=TRUE)
# global recode on marital status 
pri_data_pram_age_mar$marital_status <- mapvalues(pri_data_pram_age_mar$marital_status, from=c("Divorced", "Married/separated", "Never married", "Widowed"), to=c("Not married", "Married", "Not married", "Not married"))

# creating object 
sdc_obj <- createSdcObj(dat = pri_data_pram_age_mar, keyVars = c("age_group", "sex", "citizenship", "education", "marital_status", "zip"))

sdc_obj_k <- createSdcObj(dat = pri_data_pram_age_mar, keyVars = c("age_group", "citizenship", "marital_status", "zip"))
print(sdc_obj_k, type="kAnon")
print(sdc_obj_k, "risk")
print(sdc_obj_k@risk$global$risk)


# pram on sex variable 
sdc_obj <- pram(sdc_obj, variables = "sex", strata_variables = NULL, pd = 0.7, alpha = 0.5)

anonymized_data2 <- extractManipData(sdc_obj)

# delete the name 
anon_data_pram_global_sex_age <- anonymized_data2[, !(colnames(anonymized_data2) %in% c("dob", "age"))]
print(anon_data_pram_global_sex_age)

#write.xlsx(anon_data_pram_global_sex_age, file = "anon_data/anon_data_global2_pram.xlsx")

```
Global recoding on age
Global recoding on marital status 
```{r}
#global recode on age 
pri_data_age_mar <- pri_data
pri_data_age_mar$age_group <- cut(pri_data_age_mar$age, breaks=c(18, 30, 38, 48, 58, 70, 101), 
                    right=TRUE)
# global recode on marital status 
pri_data_age_mar$marital_status <- mapvalues(pri_data_age_mar$marital_status, from=c("Divorced", "Married/separated", "Never married", "Widowed"), to=c("Not married", "Married", "Not married", "Not married"))

# creating object 
sdc_obj <- createSdcObj(dat = pri_data_age_mar, keyVars = c("age_group", "sex", "citizenship", "education", "marital_status", "zip"))

sdc_obj_k <- createSdcObj(dat = pri_data_age_mar, keyVars = c("age_group", "citizenship", "marital_status", "zip"))
print(sdc_obj_k, type="kAnon")
print(sdc_obj_k, "risk")

anonymized_data3 <- extractManipData(sdc_obj)

# delete the dob and age  
anon_data_pram_global_sex_age <- anonymized_data3[, !(colnames(anonymized_data) %in% c("dob", "age"))]
print(anon_data_pram_global_sex_age)

#write.xlsx(anon_data_pram_global_sex_age, file = "anon_data/anon_data_global2.xlsx")

```

Global recoding on age, global recoding on marital status
pram on sex
local supression to achieve k-anonymity 
```{r}
#global recode on age 
pri_data_pram_age_mar_supression <- pri_data
pri_data_pram_age_mar_supression$age_group <- cut(pri_data_pram_age_mar_supression$age, breaks=c(18, 30, 38, 48, 58, 70, 101), 
                    right=TRUE)
# global recode on marital status 
pri_data_pram_age_mar_supression$marital_status <- mapvalues(pri_data_pram_age_mar_supression$marital_status, from=c("Divorced", "Married/separated", "Never married", "Widowed"), to=c("Not married", "Married", "Not married", "Not married"))

# creating object 
sdc_obj <- createSdcObj(dat = pri_data_pram_age_mar_supression, keyVars = c("age_group", "sex", "citizenship", "education", "marital_status", "zip"))

sdc_obj_k <- createSdcObj(dat = pri_data_pram_age_mar_supression, keyVars = c("age_group", "citizenship", "marital_status", "zip"))
# local supression 
sdc_obj <- localSuppression(sdc_obj, k = 2)

# k-anonymity computation 
sdc_obj_k <- sdc_obj_k <- localSuppression(sdc_obj, k = 2)
print(sdc_obj_k, type="kAnon")
print(sdc_obj_k, "risk")

# pram on sex variable 
sdc_obj <- pram(sdc_obj, variables = "sex", strata_variables = NULL, pd = 0.7, alpha = 0.5)

anonymized_data3 <- extractManipData(sdc_obj)

# delete the name 
pri_data_pram_age_mar_supression <- anonymized_data3[, !(colnames(anonymized_data3) %in% c("dob", "age"))]
print(pri_data_pram_age_mar_supression)
#write.xlsx(pri_data_pram_age_mar_supression, file = "anon_data/anon_data_global2_pram_supression.xlsx")

```
Global recoding on age and marital status 
pram on sex 
```{r}
#global recode on age 
pri_data_pram_age_mar_supression <- pri_data
pri_data_pram_age_mar_supression$age_group <- cut(pri_data_pram_age_mar_supression$age, breaks=c(18, 30, 38, 48, 58, 70, 101), 
                    right=TRUE)
# global recode on marital status 
pri_data_pram_age_mar_supression$marital_status <- mapvalues(pri_data_pram_age_mar_supression$marital_status, from=c("Divorced", "Married/separated", "Never married", "Widowed"), to=c("Not married", "Married", "Not married", "Not married"))

# creating object 
sdc_obj <- createSdcObj(dat = pri_data_pram_age_mar_supression, keyVars = c("age_group", "sex", "citizenship", "education", "marital_status", "zip"))

sdc_obj_k <- createSdcObj(dat = pri_data_pram_age_mar_supression, keyVars = c("age_group", "citizenship", "marital_status", "zip"))


# k-anonymity computation 
sdc_obj_k <- sdc_obj_k <- localSuppression(sdc_obj, k = 2)
print(sdc_obj_k, type="kAnon")
print(sdc_obj_k, "risk")

# pram on sex variable 
sdc_obj <- pram(sdc_obj, variables = "sex", strata_variables = NULL, pd = 0.7, alpha = 0.5)

anonymized_data3 <- extractManipData(sdc_obj)

# delete the name 
pri_data_pram_age_mar_supression <- anonymized_data3[, !(colnames(anonymized_data3) %in% c("dob", "age"))]
print(pri_data_pram_age_mar_supression)
#write.xlsx(pri_data_pram_age_mar_supression, file = "anon_data/anon_data_global2_pram_supression.xlsx")

```




PRAM on sex variable 
```{r}
sdc_obj <- createSdcObj(dat = pri_data, keyVars = c("age", "sex", "citizenship", "education", "marital_status", "zip"))

sdc_obj <- pram(sdc_obj, variables = "sex", strata_variables = NULL, pd = 0.8, alpha = 0.5)
# kAnon_violations(prammed, TRUE, 2)
anonymized_data_pram <- extractManipData(sdc_obj)
print(anonymized_data_pram)
print(sdc_obj, type="kAnon")

#write.xlsx(anonymized_data_pram, file = "anon_data/anonymized_data_pram.xlsx")

```
Swapping 
```{r}
sdc_obj <- createSdcObj(dat = pri_data, keyVars = c("age", "sex", "citizenship", "education", "marital_status", "zip"))
#pri_data$marital_status <- mapvalues(pri_data$marital_status, from=c("Divorced", "Married/separated", "Never married", "Widowed"), to=c(1, 2, 3, 4))
data_swap <- rankSwap(obj = pri_data, variables = c("marital_status"))
data_swap
```

Supression 
```{r}
sdc_obj <- createSdcObj(dat = pri_data, keyVars = c("age", "sex", "citizenship", "education", "marital_status", "zip"))

suppressed <- localSuppression(sdc_obj)
print(suppressed)
```

Other ideas:
- noise to age
- suppression
- swaping (e.g. citizenship, marital status)
- get rid of zip codes and other data we simply don't use

Questions
- does "analysis"

